m0.0.1: Microxanox - ...

### 2022-01-26 inst/manuscript/manuscript_microxanox.Rmd
intermediate

---
title: "m0.0.1: Microxanox - ..."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch & Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
keywords: metadata quality; data curation; archival; long term storage; R package;
abstract: |
  1. set the context for and purpose of the work;
  2. indicate the approach and methods;
  3. outline the main results;
  4. identify the conclusions and the wider implications.

journal: 'Ecology and Evolution - Applications'
bibliography: manuscript_bibliography.bib
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
  rmarkdown::html_vignette: default
  pdf_document:
    keep_tex: yes
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{setspace}
- \doublespacing
- \usepackage{fancyhdr}
- \fancyhead{}
- \fancyfoot{}
- \fancyhead[CO,CE]{Metadata made easy - the dmdScheme approach}
- \pagestyle{fancy}
vignette: |
  %\VignetteIndexEntry{Manuscript - microxanox} 
  %\VignetteEncoding{UTF-8} 
  %\VignetteEngine{knitr::rmarkdown}
nocite: '@bush2017'
linkcolor: blue
link-citations: yes
csl: methods-in-ecology-and-evolution.csl
---

@@

-   Facilitate reproduction of the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
-   Include new functionality that allows new research questions to be answered.

We added functionality for:

@@

We also include a function for visualising results, though **please note that it is provided for convenience, and not for production of perfectly formatted print-ready graphs.**

Here (again) is the original publication: <https://www.nature.com/articles/s41467-017-00912-x>

# The Model
Based on [@Bush2017]

@@

-   Define the default simulation conditions (i.e. parameters and initial conditions).
-   Use these defaults unless other conditions are specified.
-   The function call `run_simulation()` would therefore run a simulation with default conditions.
-   The function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with the specified initial state values, and otherwise all the specified defaults.

The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model()`. This augmented version of the model published in Bush et al 2017 can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events (see below for event definition).

# Implementation, Availability and Usage
On github [microxanox github repository](https://github.com/opetchey/microxanox/)

# Use cases

>>>>

---
title: "m0.0.1: Microxanox - ..."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch & Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
keywords: metadata quality; data curation; archival; long term storage; R package;
abstract: |
  1. set the context for and purpose of the work;
  2. indicate the approach and methods;
  3. outline the main results;
  4. identify the conclusions and the wider implications.
journal: 'Ecology and Evolution - Applications'
bibliography: manuscript_bibliography.bib
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
  rmarkdown::html_vignette: default
  pdf_document:
    keep_tex: yes
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{setspace}
- \doublespacing
- \usepackage{fancyhdr}
- \fancyhead{}
- \fancyfoot{}
- \fancyhead[CO,CE]{Metadata made easy - the dmdScheme approach}
- \pagestyle{fancy}
vignette: |
  %\VignetteIndexEntry{Manuscript - microxanox} 
  %\VignetteEncoding{UTF-8} 
  %\VignetteEngine{knitr::rmarkdown}
nocite: '@bush2017'
linkcolor: blue
link-citations: yes
csl: methods-in-ecology-and-evolution.csl
---

@@

-   Facilitate reproduction of the results of [@bush2017].
-   Include new functionality that allows new research questions to be answered.

In addition to the model described in [@bush2017], we added functionality for:

@@

The package also ingcludes some functions to viaualise the results which can function as template functions for developing graphs for own rquirements.

# The Model
The model is based on [@bush2017]. We use in this package the notation from the Table S1 of the Supplementary information of [@bush2017]), and not the notation in the main manuscript of [@bush2017]. (There is only one small difference between these: In the main text the notation is simplified to `y_S_PB` ($y_{S,PB}$) (which is `y_SR_PB` ($y_{SR,PB}$) in Table S1) and `y_S_SB` ($y_{S,SB}$) (which is y_SO_SB ($y_{SO,SB}$) in Table S1). We also use `Pr_cB` ($Pr_B$) instead of `P_CB` ($P_{CB}$).

## Model dimensions

-   Time: hours
-   Volume: litres
-   Substrate quantity: micromoles
-   Organism quantity: cells

## Modelling framework

@@

1. Define a parameter set of class `runsim_parameter` (i.e. parameters and initial conditions) by using the functions `new_runsim_parameter()`, `new_initial_state()` and `new_strain_parameter()`.
2. The function call `run_simulation(parameter)` which will run a simulation using the parameter as defined in the object `parameter`.

The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model()`. This augmented version of the model published in [@bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.

# Implementation, Availability and Usage
The package is available in a github repo at https://github.com/UZH-PEG/microxanox/.

# Use cases
This package has been used in the production of the paper [@REF_NEEDED]. It can be used for similar studies in this study systems. It is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@bush2017] and to extend it to our needs. Consequently, any more substantial, are likely to need a change in the source code.
Nevertheless, the model is structured in a way which builds on a udular structuere, so that e.g. the event defuinition can easily be changed.
